---
title: "autonomous bus vs special needs"
author: "cm"
date: "5/25/2020"
output: html_document
---
```{r}
library(MASS)
library(readr)
library(tidyverse)
library(ggplot2)
library(openxlsx)
```
```{r}
mycolor <- c("#fd8d3c", "#b30000", "#fdd0a2", "#bdbdbd","#addd8e","#006837")
```

#reading in data
```{r}
#DR
Ionia <- read_csv("../data/dr/Ionia-DR Survey 2019_April 6, 2019_09.28.csv")
Allegan <- read_csv("../data/dr/Allegan-DR Survey 2019_April 6, 2019_09.29.csv")
Flint_DR <- read_csv("../data/dr/MTA Flint-DR Survey 2019_April 6, 2019_09.30.csv")
Jackson_DR <- read_csv("../data/dr/JATA Jackson-DR Survey 2019_April 24, 2019_12.43.csv")
Adrian <- read_csv("../data/dr/Adrian -DR Survey 2019_September 17, 2019_12.03.csv")
Berrien <- read_csv("../data/dr/Berrien-DR Survey 2019_September 17, 2019_12.04.csv")
Big_Rapids <- read_csv("../data/dr/Big Rapids - DR Survey 2019_September 17, 2019_12.05.csv")
Caro <- read_csv("../data/dr/Caro -DR Survey 2019_September 17, 2019_12.05.csv")
Charlevoix <- read_csv("../data/dr/Charlevoix - DR Survey 2019_September 17, 2019_12.06.csv")
Clare <- read_csv("../data/dr/Clare -DR Survey 2019_September 17, 2019_12.06.csv")
Crawford <- read_csv("../data/dr/Crawford-DR Survey 2019_September 17, 2019_12.07.csv")
Greenville <- read_csv("../data/dr/Greenville -DR Survey 2019_September 17, 2019_12.08.csv")
Hillsdale <- read_csv("../data/dr/Hillsdale-DR Survey 2019_September 17, 2019_12.09.csv")
Kalkaska <- read_csv("../data/dr/Kalkaska -DR Survey 2019_September 17, 2019_12.11.csv")
Macatawa <- read_csv("../data/dr/Macatawa -DR Survey 2019_September 17, 2019_12.11.csv")
Midland <- read_csv("../data/dr/Midland -DR Survey 2019_September 17, 2019_12.13.csv")
Niles <- read_csv("../data/dr/Niles-DR Survey 2019_September 17, 2019_12.14.csv")
St_Joseph <- read_csv("../data/dr/St Joseph -DR Survey 2019_September 17, 2019_12.15.csv")
Twin_Cities <- read_csv("../data/dr/Twin Cities -DR Survey 2019_September 17, 2019_12.16.csv")
#---
Isabella_DR <- read_csv("../data/dr/Isabella+County-DR+Survey+2019_May+26,+2020_13.29.csv")
Marshall <- read_csv("../data/dr/Marshall-DR+Survey_May+26,+2020_14.05.csv")
Belding <- read_csv("../data/dr/Belding-DR+Survey_May+26,+2020_14.12.csv")
Clinton <- read_csv("../data/dr/Clinton+Area-DR+Survey+2019_May+26,+2020_14.17.csv")
Alma <- read_csv("../data/dr/Alma-DR+Survey_May+26,+2020_14.30.csv")
Eaton <- read_csv("../data/dr/Eaton+County-DR+Survey_May+26,+2020_14.36.csv")

#FR
Interurban_FR <- read_csv("../data/fr/Interurban -FR Survey 2019_September 20, 2019_14.44.csv")
Jackson_FR <- read_csv("../data/fr/JATA Jackson-FR Survey 2019_September 20, 2019_14.45.csv")
Macatawa_FR <- read_csv("../data/fr/Macatawa -FR Survey 2019_September 20, 2019_14.46.csv")
Flint_FR <- read_csv("../data/fr/MTA Flint-FR Survey 2019_September 20, 2019_14.47.csv")
Niles_FR <- read_csv("../data/fr/Niles-FR Survey 2019_September 20, 2019_14.48.csv")
Twin_Cities_FR <- read_csv("../data/fr/Twin Cities -FR Survey 2019_September 20, 2019_14.49.csv")
#---
Isabella_FR <- read_csv("../data/fr/Isabella-FR+Survey_May+27,+2020_11.57.csv")
```
#cleaning
```{r}
#remove Qualtrics headings
Ionia <- Ionia [-c(1,2),]
Allegan <- Allegan [-c(1,2),]
Flint_DR <- Flint_DR [-c(1,2),]
Jackson_DR <- Jackson_DR [-c(1,2),]
Adrian <- Adrian [-c(1,2),]
Berrien <- Berrien [-c(1,2),]
Big_Rapids <- Big_Rapids [-c(1,2),]
Caro <- Caro [-c(1,2),]
Charlevoix <- Charlevoix [-c(1,2),]
Clare <- Clare [-c(1,2),]
Crawford <- Crawford [-c(1,2),]
Greenville <- Greenville [-c(1,2),]
Hillsdale <- Hillsdale [-c(1,2),]
Kalkaska <- Kalkaska [-c(1,2),]
Macatawa <- Macatawa [-c(1,2),]
Midland <- Midland [-c(1,2),]
Niles <- Niles [-c(1,2),]
St_Joseph <- St_Joseph [-c(1,2),]
Twin_Cities <- Twin_Cities [-c(1,2),]
Isabella_DR <- Isabella_DR[-c(1,2),]
Marshall <- Marshall[-c(1,2),]
Belding <- Belding[-c(1,2),]
Clinton <- Clinton[-c(1,2),]
Alma <- Alma[-c(1,2),]
Eaton <- Eaton[-c(1,2),]
#---
Interurban_FR <- Interurban_FR[-c(1,2),]
Jackson_FR <- Jackson_FR[-c(1,2),]
Macatawa_FR <- Macatawa_FR[-c(1,2),]
Flint_FR <- Flint_FR[-c(1,2),]
Niles_FR <- Niles_FR[-c(1,2),]
Twin_Cities_FR <- Twin_Cities_FR[-c(1,2),]
Isabella_FR <- Isabella_FR[-c(1,2),]

#combine the first part of DR (with same formatting) into one
DR1 <- bind_rows(Ionia, Allegan, Flint_DR, Jackson_DR, Adrian, Berrien, Big_Rapids, Caro, Charlevoix, Clare, Crawford, Greenville, Hillsdale, Kalkaska, Macatawa, Midland, Niles, St_Joseph, Twin_Cities, Marshall, Belding, Clinton, Alma, Eaton) %>% select(Q17, Q18, Q19, Q22, Q25, Q26, Q34, Q35, Q36, Q27, Q28, Q29, Q30, Q31, Q32, Q32_12_TEXT, Q33)
#write.xlsx(DR1, file = "DR1.xlsx")

#Isabella_DR's question numbers were different from DR1
DR2 <- Isabella_DR %>% select(Q17, Q18, Q19, Q22, Q25, Q26, Q32, Q33, Q34, Q35, Q36, Q37, Q38, Q39, Q40, Q40_12_TEXT, Q41)

#combine all FR into one
FR1 <- bind_rows(Interurban_FR, Jackson_FR, Macatawa_FR, Flint_FR, Niles_FR, Twin_Cities_FR, Isabella_FR) %>% select(Q15, Q16, Q17, Q18, Q21, Q22, Q31,Q32,Q33,Q23,Q24,Q25,Q26,Q27,Q28,Q28_6_TEXT,Q29)
#write.xlsx(FR1, file = "FR.xlsx")
```

##helpful
```{r}
#helpfulness of the driver: 1=very dissatisfied, 5= very satisfied
DR1$Q17[DR1$Q17 == 6] <- NA
DR1 <- DR1 %>% mutate(helpful=as.numeric(Q17))
DR1$helpful <- recode(DR1$helpful, `1`=5, `2`=4, `3`=3, `4`=2, `5`=1)

DR2$Q17[DR2$Q17 == 6] <- NA
DR2 <- DR2 %>% mutate(helpful=as.numeric(Q17))
DR2$helpful <- recode(DR2$helpful, `1`=5, `2`=4, `3`=3, `4`=2, `5`=1)

FR1$Q15[FR1$Q15 == 6] <- NA
FR1 <- FR1 %>% mutate(helpful=as.numeric(Q15))
FR1$helpful <- recode(FR1$helpful, `1`=5, `2`=4, `3`=3, `4`=2, `5`=1)

```

##prof
```{r}
#professionalism of the driver: 1=very dissatisfied, 5= very satisfied
DR1$Q18[DR1$Q18 == 6] <- NA
DR1 <- DR1 %>% mutate(prof=as.numeric(Q18))
DR1$prof <- recode(DR1$prof, `1`=5, `2`=4, `3`=3, `4`=2, `5`=1)

DR2$Q18[DR2$Q18 == 6] <- NA
DR2 <- DR2 %>% mutate(prof=as.numeric(Q18))
DR2$prof <- recode(DR2$prof, `1`=5, `2`=4, `3`=3, `4`=2, `5`=1)

FR1$Q16[FR1$Q16 == 6] <- NA
FR1 <- FR1 %>% mutate(prof=as.numeric(Q16))
FR1$prof <- recode(FR1$prof, `1`=5, `2`=4, `3`=3, `4`=2, `5`=1)
```

##skill
```{r}
#driving skill of the driver: 1=very dissatisfied, 5= very satisfied
DR1$Q19[DR1$Q19 == 6] <- NA
DR1 <- DR1 %>% mutate(skill=as.numeric(Q19))
DR1$skill <- recode(DR1$skill, `1`=5, `2`=4, `3`=3, `4`=2, `5`=1)

DR2$Q19[DR2$Q19 == 6] <- NA
DR2 <- DR2 %>% mutate(skill=as.numeric(Q19))
DR2$skill <- recode(DR2$skill, `1`=5, `2`=4, `3`=3, `4`=2, `5`=1)

FR1$Q17[FR1$Q17 == 6] <- NA
FR1 <- FR1 %>% mutate(skill=as.numeric(Q17))
FR1$skill <- recode(FR1$skill, `1`=5, `2`=4, `3`=3, `4`=2, `5`=1)
```

##service
```{r}
#driving skill of the driver: 1=very dissatisfied, 5= very satisfied
DR1$Q22[DR1$Q22 == 6] <- NA
DR1 <- DR1 %>% mutate(service=as.numeric(Q22))
DR1$service <- recode(DR1$service, `1`=5, `2`=4, `3`=3, `4`=2, `5`=1)

DR2$Q22[DR2$Q22 == 6] <- NA
DR2 <- DR2 %>% mutate(service=as.numeric(Q22))
DR2$service <- recode(DR2$service, `1`=5, `2`=4, `3`=3, `4`=2, `5`=1)

FR1$Q18[FR1$Q18 == 6] <- NA
FR1 <- FR1 %>% mutate(service=as.numeric(Q18))
FR1$service <- recode(FR1$service, `1`=5, `2`=4, `3`=3, `4`=2, `5`=1)
```

##work
```{r}
#work purpose: 1=yes, 0=no
#Q25 multiple selection
DR1 <- DR1 %>% mutate(work = ifelse(grepl("1", Q25), 1, 0)) # code "work" to 1, else 0
DR1$work[is.na(DR1$Q25) == TRUE] <- NA

DR2 <- DR2 %>% mutate(work = ifelse(grepl("1", Q25), 1, 0)) # code "work" to 1, else 0
DR2$work[is.na(DR2$Q25) == TRUE] <- NA

FR1 <- FR1 %>% mutate(work = ifelse(grepl("1", Q21), 1, 0)) # code "work" to 1, else 0
FR1$work[is.na(FR1$Q21) == TRUE] <- NA
```

##med
```{r}
#medical purpose: 1=yes, 0=no
#Q25 multiple selection
DR1 <- DR1 %>% mutate(med = ifelse(grepl("2", Q25), 1, 0)) # code "medical" to 1, else 0
DR1$med[is.na(DR1$Q25) == TRUE] <- NA

DR2 <- DR2 %>% mutate(med = ifelse(grepl("2", Q25), 1, 0)) # code "medical" to 1, else 0
DR2$med[is.na(DR2$Q25) == TRUE] <- NA

FR1 <- FR1 %>% mutate(med = ifelse(grepl("2", Q21), 1, 0)) # code "medical" to 1, else 0
FR1$med[is.na(FR1$Q21) == TRUE] <- NA
```

##alt
```{r}
#alternative means: 1=yes, 0=no
#Q26 multiple selection
DR1 <- DR1 %>% mutate(alt = ifelse(grepl("1", Q26), 0, 1)) # code "not make this trip" to 0, else 1
DR1$alt[is.na(DR1$Q26) == TRUE] <- NA

DR2 <- DR2 %>% mutate(alt = ifelse(grepl("1", Q26), 0, 1)) # code "not make this trip" to 0, else 1
DR2$alt[is.na(DR2$Q26) == TRUE] <- NA

FR1 <- FR1 %>% mutate(alt = ifelse(grepl("1", Q22), 0, 1)) # code "not make this trip" to 0, else 1
FR1$alt[is.na(FR1$Q22) == TRUE] <- NA
```

##feel
```{r}
#how do you feel about AVs
DR1 <- DR1 %>% mutate(feel=Q34)

DR2 <- DR2 %>% mutate(feel=Q32)

FR1 <- FR1 %>% mutate(feel=Q31)
```

##wtr
```{r}
#willingness to ride: 1= yes, 0=no
DR1 <- DR1 %>% mutate(wtr=case_when(Q35==1 ~1, Q35==2~0)) # code "no" from 2 to 0

DR2 <- DR2 %>% mutate(wtr=case_when(Q33==1 ~1, Q35==2~0)) # code "no" from 2 to 0

FR1 <- FR1 %>% mutate(wtr=case_when(Q32==1 ~1, Q32==4~0)) # code "no" from 4 to 0
```

##concern
```{r}
#if not willing, what would be the main concerns
DR1 <- DR1 %>% mutate(concern=Q36)

DR2 <- DR2 %>% mutate(concern=Q34)

FR1 <- FR1 %>% mutate(concern=Q33)
```

##emp
```{r}
#employment: 1=in the labor force, 0=not in the labor force
DR1 <- DR1 %>% mutate(emp = ifelse(Q27 < 3, 1, 0)) # code "employed for pay" and "self-employed" to 1, else 0
DR1$emp[is.na(DR1$Q27) == TRUE] <- NA

DR2 <- DR2 %>% mutate(emp = ifelse(Q35 < 3, 1, 0)) # code "employed for pay" and "self-employed" to 1, else 0
DR2$emp[is.na(DR2$Q35) == TRUE] <- NA

FR1 <- FR1 %>% mutate(emp = ifelse(Q23 < 3, 1, 0)) # code "employed for pay" and "self-employed" to 1, else 0
FR1$emp[is.na(FR1$Q23) == TRUE] <- NA
```

##gender
```{r}
#gender: 0=male, 1=female
DR1$Q28 [DR1$Q28 == 3] <- NA
DR1 <- DR1 %>% mutate(gender=case_when(Q28==1 ~0, Q28==2~1)) # code male to 0, female to 1

DR2$Q36 [DR2$Q36 == 3] <- NA
DR2 <- DR2 %>% mutate(gender=case_when(Q36==1 ~0, Q36==2~1)) # code male to 0, female to 1

FR1$Q24 [FR1$Q24 == 3] <- NA
FR1 <- FR1 %>% mutate(gender=case_when(Q24==1 ~0, Q24==2~1)) # code male to 0, female to 1
```

##age
```{r}
#age:1=18-24, 2=25=34, 3=35-54, 4=55-64, 5=65-74, 6=75-84, 7=85 and older
DR1 <- DR1 %>% mutate(age=as.numeric(Q29))
DR1$age[DR1$age == 8] <- 3 # recode "45 to 54" from 8 to 3("35-54")

DR2 <- DR2 %>% mutate(age=as.numeric(Q37))
DR2$age[DR2$age == 8] <- 3 # recode "45 to 54" from 8 to 3("35-54")

FR1 <- FR1 %>% mutate(age=as.numeric(Q25))
FR1$age[FR1$age == 14] <- 7 # recode "85 and older" from 14 to 7
FR1$age[FR1$age == 8] <- 3 # recode "45 to 54" from 8 to 3("35-54")
```

##income
```{r}
DR1 <- DR1 %>% mutate(income=as.numeric(Q30))

DR2 <- DR2 %>% mutate(income=as.numeric(Q38))

FR1 <- FR1 %>% mutate(income=as.numeric(Q26))
```

##race & racecat
```{r}
#race: 1=black, 2=asian, 3=white, 4=hispanic, 5=native american, 6=pacific islander, 7=other
#racecat: 1=black, 3=white, 7=other (including hispanic)
DR1 <- DR1 %>% mutate(race=as.numeric(Q31))
DR1 <- DR1 %>% mutate(racecat=case_when(race==1~1, race==2~7, race==3~3, race==4~7, race==5~7,race==6~7,race==7~7))

DR2 <- DR2 %>% mutate(race=as.numeric(Q39))
DR2 <- DR2 %>% mutate(racecat=case_when(race==1~1, race==2~7, race==3~3, race==4~7, race==5~7,race==6~7,race==7~7))

FR1 <- FR1 %>% mutate(race=as.numeric(Q27))
FR1 <- FR1 %>% mutate(racecat=case_when(race==1~1, race==2~7, race==3~3, race==4~7, race==5~7,race==6~7,race==7~7))
```

##accom
```{r}
#special needs: 1=yes, 0=no
DR1 <- DR1 %>% mutate(accom = ifelse(Q32==7, 0, 1)) # code only select "no special needs" to 0, else 1
DR1$accom[is.na(DR1$Q32) == TRUE] <- NA

DR2 <- DR2 %>% mutate(accom = ifelse(Q40==7, 0, 1)) # code only select "no special needs" to 0, else 1
DR2$accom[is.na(DR2$Q40) == TRUE] <- NA

FR1 <- FR1 %>% mutate(accom = ifelse(Q28==1, 0, 1)) # code only select "no special needs" to 0, else 1
FR1$accom[is.na(FR1$Q28) == TRUE] <- NA
```

##blind
```{r}
#blindness: 1=yes, 0=no
DR1 <- DR1 %>% mutate(blind = ifelse(grepl("8", Q32), 1, 0)) # code "blindness/ visual impairment" to 1, else 0
DR1$blind[is.na(DR1$Q32) == TRUE] <- NA

DR2 <- DR2 %>% mutate(blind = ifelse(grepl("8", Q40), 1, 0)) # code "blindness/ visual impairment" to 1, else 0
DR2$blind[is.na(DR2$Q40) == TRUE] <- NA

FR1 <- FR1 %>% mutate(blind = ifelse(grepl("2", Q28), 1, 0)) # code "blindness/ visual impairment" to 1, else 0
FR1$blind[is.na(FR1$Q28) == TRUE] <- NA
```

##deaf
```{r}
#deaf: 1=yes, 0=no
DR1 <- DR1 %>% mutate(deaf = ifelse(grepl("9", Q32), 1, 0)) # code "deaf/hard of hearing" to 1, else 0
DR1$deaf[is.na(DR1$Q32) == TRUE] <- NA

DR2 <- DR2 %>% mutate(deaf = ifelse(grepl("9", Q40), 1, 0)) # code "deaf/hard of hearing" to 1, else 0
DR2$deaf[is.na(DR2$Q40) == TRUE] <- NA

FR1 <- FR1 %>% mutate(deaf = ifelse(grepl("3", Q28), 1, 0)) # code "deaf/hard of hearing" to 1, else 0
FR1$deaf[is.na(FR1$Q28) == TRUE] <- NA
```

##mobility
```{r}
#mobility disability: 1=yes, 0=no
DR1 <- DR1 %>% mutate(mobility = ifelse(grepl("10", Q32), 1, 0)) # code "mobility disabilities" to 1, else 0
DR1$mobility[is.na(DR1$Q32) == TRUE] <- NA

DR2 <- DR2 %>% mutate(mobility = ifelse(grepl("10", Q40), 1, 0)) # code "mobility disabilities" to 1, else 0
DR2$mobility[is.na(DR2$Q40) == TRUE] <- NA

FR1 <- FR1 %>% mutate(mobility = ifelse(grepl("4", Q28), 1, 0)) # code "mobility disabilities" to 1, else 0
FR1$mobility[is.na(FR1$Q28) == TRUE] <- NA
```

##psy
```{r}
#psychiatric disability: 1=yes, 0=no
DR1 <- DR1 %>% mutate(psy = ifelse(grepl("11", Q32), 1, 0)) # code "psychiatric disabilities" to 1, else 0
DR1$psy[is.na(DR1$Q32) == TRUE] <- NA

DR2 <- DR2 %>% mutate(psy = ifelse(grepl("11", Q40), 1, 0)) # code "psychiatric disabilities" to 1, else 0
DR2$psy[is.na(DR2$Q40) == TRUE] <- NA

FR1 <- FR1 %>% mutate(psy = ifelse(grepl("5", Q28), 1, 0)) # code "psychiatric disabilities" to 1, else 0
FR1$psy[is.na(FR1$Q28) == TRUE] <- NA
```
##otherdis
```{r}
#other disability: 1=yes, 0=no
DR1 <- DR1 %>% mutate(otherdis = ifelse(grepl("12", Q32), 1, 0)) # code "other disabilities" to 1, else 0
DR1$otherdis[is.na(DR1$Q32) == TRUE] <- NA

DR2 <- DR2 %>% mutate(otherdis = ifelse(grepl("12", Q40), 1, 0)) # code "other disabilities" to 1, else 0
DR2$otherdis[is.na(DR2$Q40) == TRUE] <- NA

FR1 <- FR1 %>% mutate(otherdis = ifelse(grepl("6", Q28), 1, 0)) # code "other disabilities" to 1, else 0
FR1$otherdis[is.na(FR1$Q28) == TRUE] <- NA
```

```{r}
#examine other disability
#DR1 %>% group_by(Q32_12_TEXT) %>% count()
#DR2 %>% group_by(Q40_12_TEXT) %>% count()
#FR1 %>% group_by(Q28_6_TEXT) %>% count()
```
```{r}
#manually code "other"
DR1$mobility[DR1$Q32_12_TEXT=="Cane"] <- 1
DR1$otherdis[DR1$Q32_12_TEXT=="Cane"] <- 0
DR1$mobility[DR1$Q32_12_TEXT=="for 1 month was using a walker"] <- 1
DR1$otherdis[DR1$Q32_12_TEXT=="for 1 month was using a walker"] <- 0
DR1$accom[DR1$Q32_12_TEXT=="I don't require any accommodations"] <- 0
DR1$otherdis[DR1$Q32_12_TEXT=="I don't require any accommodations"] <- 0
DR1$mobility[DR1$Q32_12_TEXT=="Lift on occasion"] <- 1
DR1$otherdis[DR1$Q32_12_TEXT=="Lift on occasion"] <- 0
DR1$mobility[DR1$Q32_12_TEXT=="Mobility"] <- 1
DR1$otherdis[DR1$Q32_12_TEXT=="Mobility"] <- 0
DR1$accom[DR1$Q32_12_TEXT=="No"] <- 0
DR1$otherdis[DR1$Q32_12_TEXT=="No"] <- 0
DR1$accom[DR1$Q32_12_TEXT=="none"] <- 0
DR1$otherdis[DR1$Q32_12_TEXT=="none"] <- 0
DR1$accom[DR1$Q32_12_TEXT=="None"] <- 0
DR1$otherdis[DR1$Q32_12_TEXT=="None"] <- 0
DR1$mobility[DR1$Q32_12_TEXT=="wheelchaie"] <- 1
DR1$otherdis[DR1$Q32_12_TEXT=="wheelchaie"] <- 0

DR2$mobility[DR2$Q40_12_TEXT=="use walker and wagon"] <- 1
DR2$otherdis[DR2$Q40_12_TEXT=="use walker and wagon"] <- 0

FR1$accom[FR1$Q28_6_TEXT=="N/A"] <- 0
FR1$otherdis[FR1$Q28_6_TEXT=="N/A"] <- 0
FR1$accom[FR1$Q28_6_TEXT=="none"] <- 0
FR1$otherdis[FR1$Q28_6_TEXT=="none"] <- 0
FR1$accom[FR1$Q28_6_TEXT=="None"] <- 0
FR1$otherdis[FR1$Q28_6_TEXT=="None"] <- 0
FR1$psy[FR1$Q28_6_TEXT=="Psychdisabilities"] <- 1
FR1$otherdis[FR1$Q28_6_TEXT=="Psychdisabilities"] <- 0
FR1$mobility[FR1$Q28_6_TEXT=="Walker"] <- 1
FR1$otherdis[FR1$Q28_6_TEXT=="Walker"] <- 0
FR1$mobility[FR1$Q28_6_TEXT=="Walking cane"] <- 1
FR1$otherdis[FR1$Q28_6_TEXT=="Walking cane"] <- 0
FR1$mobility[FR1$Q28_6_TEXT=="walking disability"] <- 1
FR1$otherdis[FR1$Q28_6_TEXT=="walking disability"] <- 0
FR1$mobility[FR1$Q28_6_TEXT=="Wheelchair"] <- 1
FR1$otherdis[FR1$Q28_6_TEXT=="Wheelchair"] <- 0
```
##comment
```{r}
#any other comments about the transportation service
DR1 <- DR1 %>% mutate(comment=Q33)

DR2 <- DR2 %>% mutate(comment=Q41)

FR1 <- FR1 %>% mutate(comment=Q29)
```
```{r}
#cleaned version
dr1 <- DR1 %>% select(helpful, prof, skill, service, work, med, alt, feel, wtr, concern, emp, gender, age, income, race, racecat, accom, blind, deaf, mobility, psy, otherdis, comment)
dr1 <- dr1[rowSums(is.na(dr1)) != ncol(dr1), ] # remove rows if all NA

dr2 <- DR2 %>% select(helpful, prof, skill, service, work, med, alt, feel, wtr, concern, emp, gender, age, income, race, racecat, accom, blind, deaf, mobility, psy, otherdis, comment)
dr2 <- dr2[rowSums(is.na(dr2)) != ncol(dr2), ] # remove rows if all NA

dr <- bind_rows(dr1, dr2) %>% mutate(type ="DR") # combine dr1 and dr2 into dr

fr <- FR1 %>% select(helpful, prof, skill, service, work, med, alt, feel, wtr, concern, emp, gender, age, income, race, racecat, accom, blind, deaf, mobility, psy, otherdis, comment)
fr <- fr[rowSums(is.na(fr)) != ncol(fr), ] # remove rows if all NA
fr <- fr %>% mutate(type="FR")
```

```{r}
all <- bind_rows(dr, fr)
#glimpse(all)
```

```{r}
#d<- all %>% drop_na(wtr)
d <- all[!(is.na(all$wtr)==TRUE&is.na(all$feel)==TRUE),] # remove rows if both wtr and feel (dependent v.) are NA
#glimpse(d)
#write.xlsx(d, file = "av_special_needs.xlsx")
```

##dis
```{r}
#disability combined: 1=no special needs, 2= blind only, 3=deaf only, 4=mobility only, 5= psychiatric only, 6= otherdisability only, 7= more than one disability
d <- d %>% mutate(dis=case_when(accom==0~1, blind==1&deaf==0&mobility==0&psy==0&otherdis==0~2, blind==0&deaf==1&mobility==0&psy==0&otherdis==0~3, blind==0&deaf==0&mobility==1&psy==0&otherdis==0~4, blind==0&deaf==0&mobility==0&psy==1&otherdis==0~5, blind==0&deaf==0&mobility==0&psy==0&otherdis==1~6))
d$dis[is.na(d$dis)==TRUE] <- 7 #code both NA and multi-disability to 7
d$dis[is.na(d$accom) == TRUE] <- NA # code NA to NA so 7=multi-disability
```
##senior
```{r}
#senior: 1= yes, 0=no
d <- d %>% mutate(senior=case_when(age<5~0, age>=5~1)) # code 65 and older as seniors
```
##medrely
```{r}
#rely on PT for med: 1= yes, 0=no
d <- d %>% mutate(medrely=ifelse(med==1&alt==0, 1, 0))
d$medrely[is.na(d$med) == TRUE] <- NA
d$medrely[is.na(d$alt) == TRUE] <- NA
```

##workrely
```{r}
#rely on PT for work: 1= yes, 0=no
d <- d %>% mutate(workrely=ifelse(work==1&alt==0, 1, 0))
d$medrely[is.na(d$work) == TRUE] <- NA
d$medrely[is.na(d$alt) == TRUE] <- NA
```
```{r}
#write.xlsx(d, file = "av_special_needs_withdis.xlsx")
```

#------------
#open-ends coding

```{r}
d$index <- seq.int(nrow(d)) # add index to master sheet
openend <- read_csv("av_special_needs_openends_coded.csv")
d <- left_join(d, openend)
```

##feel_senti
```{r}
#feel: 1= negative, 2=nutural, 3=positive
d <- d %>% mutate(feel_senti=case_when(feel_code_td==1~1, feel_code_td==2~1, feel_code_td==3~1,feel_code_td==4~1, feel_code_td==5~2,feel_code_td==6~3, feel_code_td==7~3, feel_code_td==0~0)) 
d$feel_senti[d$feel_senti==0] <- NA
d %>% group_by(feel_senti) %>% count()
```
##feel_less
```{r}
# feel less codes: 1=negative simple response, 2= safety concern, 3=cost concern, 4=neutral, 5= positive simple response, 6=benefits
d <- d %>% mutate(feel_less=case_when(feel_code_td==1~1, feel_code_td==2~2, feel_code_td==3~2,feel_code_td==4~3, feel_code_td==5~4,feel_code_td==6~5, feel_code_td==7~6, feel_code_td==0~0))
d$feel_less[d$feel_less==0] <- NA
d %>% group_by(feel_less) %>% count()
```

##concern_human
```{r}
#not willing to ride because of prefer human driver, interaction
d <- d %>% mutate(concern_human=ifelse(concern_code_more==6, 1, 0)) 
d$concern_human[is.na(d$concern_code_more) == TRUE] <- NA
```

##concern_safety
```{r}
#not willing to ride because of safety concerns, including tech concerns
d <- d %>% mutate(concern_safety=ifelse(concern_code_more==3|concern_code_more==2|concern_code_more==4, 1, 0)) 
d$concern_safety[is.na(d$concern_code_more) == TRUE] <- NA
```
##concern_tech
```{r}
#not willing to ride because of technological fears
d <- d %>% mutate(concern_tech=ifelse(concern_code_more==2|concern_code_more==4|concern_code_more==8, 1, 0)) 
d$concern_tech[is.na(d$concern_code_more) == TRUE] <- NA
```
##concern_other
```{r}
#not willing to ride for other reasons
d <- d %>% mutate(concern_other=ifelse(concern_code_more==1|concern_code_more==5|concern_code_more==7|concern_code_more==9|concern_code_more==11|concern_code_more==16, 1, 0)) 
d$concern_other[is.na(d$concern_code_more) == TRUE] <- NA
```
##concern_blank
```{r}
#not willing to ride but did not give reasons
d <- d %>% mutate(concern_blank=ifelse(concern_code_more==0, 1, 0)) 
d$concern_blank[is.na(d$concern_code_more) == TRUE] <- NA
```

##wtrno
```{r}
# a subset of not willing to ride
wtrno <- d %>% filter(wtr==0)
wtrno$concern_blank[is.na(wtrno$concern_blank)==T] <- 1 # one mismach blank
wtrno$concern_human[is.na(wtrno$concern_human)==T] <- 0
wtrno$concern_safety[is.na(wtrno$concern_safety)==T] <- 0
wtrno$concern_tech[is.na(wtrno$concern_tech)==T] <- 0
wtrno$concern_other[is.na(wtrno$concern_other)==T] <- 0
```

#------------
#descriptives
```{r}
print(colSums(!is.na(d)))
```

```{r}
varde<- function(data,variable){
  data %>% filter(!is.na({{variable}})) %>% group_by({{variable}}) %>% summarise(n=n()) %>% mutate(freq=round((n/sum(n)*100),2))
  }
```
```{r}
varde(d, gender)
```
```{r}
varde(d, age)
```
```{r}
varde(d, income)
```

```{r}
varde(d, emp)
```
```{r}
varde(d, race)
```
```{r}
varde(d, dis)
```
```{r}
varde(d, medrely)
```
```{r}
varde(d, workrely)
```
```{r}
vardea <- function(data, accom_variable, variable){
  data %>% filter(!is.na({{variable}})) %>%filter(!is.na({{accom_variable}})) %>% group_by({{accom_variable}}, {{variable}}) %>% summarise(n=n()) %>% mutate(freq=round((n/sum(n)*100),2))
  }
```
```{r}
d %>% filter(!is.na(gender)) %>% group_by(accom,gender) %>% summarise(n=n()) %>% mutate(freq=round((n/sum(n)*100),2))
```
```{r}
vardea(d, accom, gender)
```
```{r}
vardea(d, accom, age)
```
```{r}
vardea(d, accom, income)
```
```{r}
vardea(d, accom, emp)
```

```{r}
vardea(d, accom, race)
```

#------------
#stats

```{r}
mcfadden <- function(m) { m.null <- update(m, . ~ 1); LLf <- logLik(m); LL0 <- logLik(m.null); as.vector(1-(LLf/LL0))}
coxsnell <- function(m) { m.null <- update(m, . ~ 1); LLf <- logLik(m); LL0 <- logLik(m.null); N <- length(resid(m)); as.vector(1 - exp((2/N) * (LL0 - LLf)))}
nagelkerke <- function(m) { m.null <- update(m, . ~ 1); LLf <- logLik(m); LL0 <- logLik(m.null); N <- length(resid(m)); as.vector((1 - exp((2/N) * (LL0 - LLf))) / (1 - exp(LL0)^(2/N)))}
```
##Hypothesis 1: Bus riders who have special needs are more likely to reply on public transit than those who do not.

```{r}
d <- d %>% mutate(rely= 1-alt)
```
```{r}
summary(model <- glm(rely ~ as.factor(dis) + as.factor(gender) + age + income + as.factor(emp)+ as.factor(racecat), data = d, family = binomial(logit)))
list(McFadden=mcfadden(model), CoxSnell=coxsnell(model), Nagelkerke=nagelkerke(model))
round(exp(coef(model)), 2)
nobs(model)
```
```{r}
# probabilities
plogis(1.19843) # vision impaired
plogis(0.58743) # mobility restricted
plogis(0.96339) # multiple needs
```

```{r}
d %>% filter(!is.na(alt)) %>% group_by(dis) %>% summarise(n=n(), pct_rely=round((1-mean(alt))*100, 2))
```
```{r}
# average dependent on PT
1-mean(d$alt, na.rm=T)
```
```{r}
d  %>% group_by(alt) %>% summarise(n=n())
```
```{r}
d %>% filter(!is.na(alt)) %>% group_by(accom,) %>% summarise(n=n(), pct_rely=1-mean(alt))
```
```{r}
d %>% filter(accom==0) %>% group_by(alt) %>% summarise(n=n())
```
```{r}
summary(model <- glm(rely ~ as.factor(accom) + as.factor(gender) + age + income + as.factor(emp)+ as.factor(racecat), data = d, family = binomial(logit)))
list(McFadden=mcfadden(model), CoxSnell=coxsnell(model), Nagelkerke=nagelkerke(model))
round(exp(coef(model)), 2)
nobs(model)
```

##Hypothesis 2: Bus riders who have special needs are more likely to be willing to use autonomous public transit than those who do not.

```{r}
# demographics only
summary(model <- glm(wtr ~ as.factor(dis)+ as.factor(gender) + age +income +as.factor(emp) +as.factor(racecat), data = d, family = binomial(logit)))
list(McFadden=mcfadden(model), CoxSnell=coxsnell(model), Nagelkerke=nagelkerke(model))
round(exp(coef(model)), 2)
nobs(model)
```
```{r}
# rely only
summary(model <- glm(wtr ~ as.factor(rely) , data = d, family = binomial(logit)))
list(McFadden=mcfadden(model), CoxSnell=coxsnell(model), Nagelkerke=nagelkerke(model))
round(exp(coef(model)), 2)
nobs(model)
```

```{r}
# demographics + rely
summary(model <- glm(wtr ~ as.factor(dis)+as.factor(gender) + age +income +as.factor(emp) +as.factor(racecat)+as.factor(rely), data = d, family = binomial(logit)))
list(McFadden=mcfadden(model), CoxSnell=coxsnell(model), Nagelkerke=nagelkerke(model))
round(exp(coef(model)), 2)
nobs(model)
```

```{r}
d %>% filter(!is.na(wtr)) %>% group_by(dis) %>% summarise(n=n(), pct_willing=round(mean(wtr)*100, 2))
```
```{r}
mean(d$wtr,na.rm = T)
```
```{r}
d %>% filter(accom==1) %>% summarize(mean(wtr, na.rm = T))
```
```{r}
d %>% filter(accom==0) %>% summarize(mean(wtr, na.rm = T))
```

```{r}
# demographics only
summary(model <- glm(wtr ~ as.factor(accom)+ as.factor(gender) + age +income +as.factor(emp) +as.factor(racecat), data = d, family = binomial(logit)))
list(McFadden=mcfadden(model), CoxSnell=coxsnell(model), Nagelkerke=nagelkerke(model))
round(exp(coef(model)), 2)
nobs(model)
```

##Hypothesis 3: There is no difference in AV perceptions between bus riders with and without special needs.

```{r}
summary(model<- polr(as.factor(feel_senti) ~ as.factor(dis) + as.factor(gender) +age +income+ as.factor(emp)+as.factor(racecat), Hess=T, data=d))
st <- coef(summary(model))
pval <- pnorm(abs(st[, "t value"]),lower.tail = FALSE)* 2
st <- cbind(st, "p value" = round(pval,4))
st
list(McFadden=mcfadden(model), CoxSnell=coxsnell(model), Nagelkerke=nagelkerke(model))
round(exp(coef(model)), 2)
nobs(model)
```
```{r}
exp(-1.465360184)
```
```{r}
exp(-0.409787441)
```
```{r}
summary(model<- polr(as.factor(feel_senti) ~ wtr, Hess=T, data=d))
st <- coef(summary(model))
pval <- pnorm(abs(st[, "t value"]),lower.tail = FALSE)* 2
st <- cbind(st, "p value" = round(pval,4))
st
list(McFadden=mcfadden(model), CoxSnell=coxsnell(model), Nagelkerke=nagelkerke(model))
round(exp(coef(model)), 2)
nobs(model)
```
```{r}
d %>% filter(!is.na(feel_less)) %>% filter(!is.na(dis)) %>%ggplot()+
  geom_histogram(aes(x=dis, fill=as.factor(feel_less)),position="fill", binwidth = 1)+
  scale_fill_manual(values=mycolor, labels= c("1"="Negative simple response", "2"="Safety concern", "3"="Cost concern",  "4"="Neutral or neutralized response","5"="Positive simple response", "6"="Benefits"))
```

```{r}
d %>% filter(!is.na(feel_less)) %>% group_by(dis) %>% count(feel_less) %>% mutate(freq=n/sum(n))
```

```{r}
d %>% filter(!is.na(feel_less)) %>% filter(accom==1) %>% group_by(feel_less)  %>% summarise(n=n()) %>% mutate(freq=n/sum(n))
```

```{r}
d %>% filter(!is.na(feel_less)) %>% filter(accom==1) %>% group_by(dis, feel_less)  %>% summarise(n=n()) 
```
```{r}
summary(model<- polr(as.factor(feel_senti) ~ as.factor(accom) + as.factor(gender) +age +income+ as.factor(emp)+as.factor(racecat), Hess=T, data=d))
st <- coef(summary(model))
pval <- pnorm(abs(st[, "t value"]),lower.tail = FALSE)* 2
st <- cbind(st, "p value" = round(pval,4))
st
list(McFadden=mcfadden(model), CoxSnell=coxsnell(model), Nagelkerke=nagelkerke(model))
round(exp(coef(model)), 2)
nobs(model)
```
##Hypothesis 4: There is no difference in concerns over AVs in terms of safety, absence of human interactions, and technological fear between bus riders with and without special needs
```{r}
count(wtrno)
```
```{r}
wtrno %>% filter(accom==1) %>% group_by(concern_safety) %>% summarise(n=n()) %>% mutate(freq=round(n/sum(n)*100,2))
```
```{r}
wtrno %>% filter(accom==1) %>% group_by(concern_human) %>% summarise(n=n()) %>% mutate(freq=round(n/sum(n)*100,2))
```
```{r}
wtrno %>% filter(accom==1) %>% group_by(concern_tech) %>% summarise(n=n()) %>% mutate(freq=round(n/sum(n)*100,2))
```
```{r}
wtrno %>% filter(accom==1)%>% group_by(concern_other) %>% summarise(n=n()) %>% mutate(freq=round(n/sum(n)*100,2))
```
```{r}
wtrno %>% group_by(concern_blank) %>% summarise(n=n()) %>% mutate(freq=round(n/sum(n)*100,2))
```
```{r}
wtrno %>% filter(accom==1) %>% group_by(concern_blank) %>% summarise(n=n()) %>% mutate(freq=round(n/sum(n)*100,2))
```
```{r}
#not willing to ride because of concerns over safety
summary(model <- glm(concern_safety ~ as.factor(dis) +as.factor(gender) + age +income+ as.factor(emp)+as.factor(racecat), data = wtrno, family = binomial(logit)))
list(McFadden=mcfadden(model), CoxSnell=coxsnell(model), Nagelkerke=nagelkerke(model))
round(exp(coef(model)), 2)
nobs(model)
```

```{r}
wtrno %>% filter(!is.na(concern_safety)) %>% group_by(dis) %>% summarize(n=n(), mean(concern_safety))
```

```{r}
#not willing to ride because of concerns over no human
summary(model <- glm(concern_human ~ as.factor(dis) +as.factor(gender) +age +income+ as.factor(emp)+as.factor(racecat), data = wtrno, family = binomial(logit)))
list(McFadden=mcfadden(model), CoxSnell=coxsnell(model), Nagelkerke=nagelkerke(model))
round(exp(coef(model)), 2)
nobs(model)
```

```{r}
wtrno %>% filter(!is.na(concern_human)) %>% group_by(dis) %>% summarize(n=n(), mean(concern_human))
```

```{r}
#not willing to ride because of concerns over tech
summary(model <- glm(concern_tech ~ as.factor(dis) +as.factor(gender) +age +income+ as.factor(emp)+as.factor(racecat), data = wtrno, family = binomial(logit)))
list(McFadden=mcfadden(model), CoxSnell=coxsnell(model), Nagelkerke=nagelkerke(model))
round(exp(coef(model)), 2)
nobs(model)
```

```{r}
wtrno %>% filter(!is.na(concern_tech)) %>% group_by(dis) %>% summarize(n=n(), mean(concern_tech))
```
```{r}
#not willing to ride because of concerns over safety
summary(model <- glm(concern_safety ~ as.factor(accom) +as.factor(gender) + age +income+ as.factor(emp)+as.factor(racecat), data = wtrno, family = binomial(logit)))
list(McFadden=mcfadden(model), CoxSnell=coxsnell(model), Nagelkerke=nagelkerke(model))
round(exp(coef(model)), 2)
nobs(model)
```
```{r}
#not willing to ride because of concerns over no human
summary(model <- glm(concern_human ~ as.factor(accom) +as.factor(gender) +age +income+ as.factor(emp)+as.factor(racecat), data = wtrno, family = binomial(logit)))
list(McFadden=mcfadden(model), CoxSnell=coxsnell(model), Nagelkerke=nagelkerke(model))
round(exp(coef(model)), 2)
nobs(model)
```
```{r}
#not willing to ride because of concerns over tech
summary(model <- glm(concern_tech ~ as.factor(accom) +as.factor(gender) +age +income+ as.factor(emp)+as.factor(racecat), data = wtrno, family = binomial(logit)))
list(McFadden=mcfadden(model), CoxSnell=coxsnell(model), Nagelkerke=nagelkerke(model))
round(exp(coef(model)), 2)
nobs(model)
```

#------------
#dataset to submit

```{r}
dataset <- d %>% mutate(employment=emp, special_needs=dis) %>% mutate(race2=case_when(race==1~1, race==2~2, race==3~3, race==4~6, race==5~4,race==6~5,race==7~6)) # put hispanic into other (suggested by Z)
dataset <- dataset %>% select(rely, feel, wtr, concern, employment, gender, age, income, race2, special_needs) %>% rename(race = race2)
glimpse(dataset)
write.xlsx(dataset, file = "dataset.xlsx")
```

